### 2015-11-26 paper.tex

\documentclass[11pt]{article}

\usepackage{deauthor,times,graphicx}

% JobGraph fig
\usepackage{tikz}
\usetikzlibrary{arrows,automata,positioning,shapes.geometric}

% trademark symbol
\usepackage{textcomp}

\usepackage{float}

\newcommand{\TODO}[1]{\textcolor{red}{TODO: #1}}

% code listings
\usepackage{listings}

\newcommand{\sectionautorefname}{Section}%

\newcommand{\para}[1]{\vspace{2mm}\noindent\textbf{#1}}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

% there is no built in support or Scala yet, good enough
\lstset{frame=l,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  xleftmargin=5pt,
%   framexleftmargin=-1pt,
  columns=flexible,
  basicstyle={\scriptsize\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{black},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=4,
  %for scala
  emph={%  
    object, def, val, zip, window, trigger, evict%
    },emphstyle={\color{blue}\textbf}%
}%

% url references
\usepackage{hyperref}

% in paragraph enums
\usepackage{paralist}

\usepackage{subfigure}

\def\definitionautorefname{Definition}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}
\def\algorithmautorefname{Algorithm}
\def\figureautorefname{Figure}

\begin{document}

\title{Apache Flink\texttrademark: Unified Stream and Batch Processing \\ in a Single Engine}

\author{Authors}

\maketitle

\begin{abstract}
Apache Flink is an open source system for processing streaming and batch data. Flink is built on the philosophy that many classes of data processing applications, including real-time analytics, continuous data pipelines, historic data processing (batch), and iterative algorithms (machine learning, graph analysis) can be expressed and executed as pipelined fault tolerant dataflows. In this paper, we present Flink’s architecture and expand on how a (seemingly diverse) set of use cases can be unified under the same execution model.
\end{abstract}

\input{intro}
\input{architecture}
\input{exec-model}
\input{streaming}
\input{batch}
\input{related}

\section{Conclusion}
\label{sec:conclusions}
In this paper we presented Apache Flink, a platform that implements one universal dataflow engine designed to perform both stream and batch analytics. Flink's dataflow engine treats operator state and logical intermediate results as first class citizens, and is used by the batch and a data stream APIs. The streaming API that is built on top of  Flink's streaming dataflow engine, provides the means to keep recoverable state, and to partition, transform, discretize and aggregate a data stream. While batch computations are, in theory, a special case of a streaming computations, Flink treats them specially, by optimizing their execution with a novel query optimizer, and by implementing blocking operators that gracefully spill to disk in the absence of memory. 
{
\small
\bibliographystyle{abbrv}
\bibliography{references}
}
\end{document}